Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-12-06 bayesian-analysis/kundinger-submit.tex
update.

\documentclass[12pt,letterpaper]{article}

\usepackage[hypertexnames=false]{hyperref}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{natbib}

% Custom packages
\usepackage{amsthm, amssymb}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}
\usepackage{multirow}
\usepackage{sectsty}
\usepackage{tabularx}
\usepackage{tikz}           % vector graphics
\usepackage{bm}             % bold math symbols
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{import}
\usepackage{titling}
\usetikzlibrary{arrows, backgrounds, patterns, matrix, shapes, fit, 
  calc, shadows, plotmarks}

\graphicspath{{./notes/figures/}}

% Custom commands
\let\oldvec\vec
\renewcommand\vec{\bm}
\newcommand{\simfn}{\mathtt{sim}} % similarity function
\newcommand{\truncsimfn}{\underline{\simfn}} % truncated similarity function
\newcommand{\blockfn}{\mathtt{BlockFn}} % blocking function
\newcommand{\distfn}{\mathtt{dist}} % distance function
\newcommand{\valset}{\mathcal{V}} % attribute value set
\newcommand{\entset}{\mathcal{R}} % set of records that make up an entity
\newcommand{\partset}{\mathcal{E}} % set of entities that make up a partition
\newcommand{\1}[1]{\mathbb{I}\!\left[#1\right]} % indicator function
\newcommand{\euler}{\mathrm{e}} % Euler's constant
\newcommand{\dblink}{\texttt{\upshape \lowercase{d-blink}}} % Name of scalable Bayesian ER model
\newcommand{\blink}{\texttt{\upshape \lowercase{blink}}} % Name of original Bayesian ER model
\def\spacingset#1{\renewcommand{\baselinestretch}%
  {#1}\small\normalsize} \spacingset{1}

\newtheorem{remark}{Remark}
\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

% DON'T change margins - should be 1 inch all around.
\addtolength{\oddsidemargin}{-.5in}%
\addtolength{\evensidemargin}{-.5in}%
\addtolength{\textwidth}{1in}%
\addtolength{\textheight}{1.3in}%
\addtolength{\topmargin}{-.8in}%

\sectionfont{\large\nohang\centering\MakeUppercase}

\title{Efficient and Scalable Bayesian Bipartite Matching with Fast Beta Linkage  (fabl)}
\author{Brian Kundinger\textsuperscript{a} \and
  Jerome Reiter\textsuperscript{a} \and 
  Rebecca C.~Steorts \textsuperscript{b}}
\date{
 \textsuperscript{a} Department of Statistical Science, Duke University \\
 \textsuperscript{b} Department of Statistical Science and Computer Science, Duke University\\Principal Mathematical Statistician, United States Census Bureau\\[2ex]
  \today}

\begin{document}
\maketitle

\bigskip
\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.

%In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends a recent Bayesian Fellegi Sunter model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space. This allows us to employ hashing techniques that hasten calculations and reduce the overall computational complexity.  In addition, we are able to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called \emph{storage efficient indexing}. Through simulation studies and a case study of homicides from the El Salvadoran Civil War, we show that our method has markedly increased speed with minimal loss of accuracy.
\end{abstract}

\noindent%
{\it Keywords:} record linkage, fusion, hashing, distributed computing, Fellegi Sunter

\newpage
\spacingset{1.5}

@@

%\hypertarget{scale}{%
%	\subsection{Scaling to a Larger Simulation Study}\label{scale}}
%
%To demonstrate the scale of the linkage tasks possible under \texttt{fabl}, we concatenate two sets of 40 simulated datasets from the original simulation studies of \cite{sadinle_bayesian_2017}, creating two larger files, each of 20,000 records (40,000 records total). Under standard Fellegi Sunter procedures, this would require 400,000,000 comparison vectors, each consisting for four integers, resulting in a final comparison matrix about 6.4 GB in size. Under default settings on a personal computer, \texttt{R} does not allow storing an object this size, so \texttt{BRL} is not usable on this linkage task. Through \texttt{fabl} however, using just one machine, we sequentially compare records across chunks, hash results, and then synthesize summary statistics for the entire simulation.  
% 
%For simplicity, we partition one dataset into 20 smaller chunks, and leave the second dataset fully intact. We compare records, hash results, and then synthesize summary statistics for all 20 chunk comparisons. The resulting data object is only 90 MB, about 1\% the size of the object required under the standard method. Executed sequentially, these comparisons and the Gibbs sampler take about one hour to run. However, this could be substantially sped up using distributed computing.
%
%Turning to evaluation metrics, this simulation achieved 96.5\% recall and 97.7\% precision, with an overall F-measure of 97.1\% F-measure. This is slightly worse performance than  in the smaller simulation studies; this is expected because it is naturally more difficult to link more records with the same amount of information. With more linkage fields, \texttt{fabl} maintains the high accuracy in Section~\ref{accuracy}.

@@

%\hypertarget{violations}{%
%	\subsection{Violations of One-to-One Matching}\label{violations}}
%
%As noted previously, the application exhibits many situations in which bipartite matching is difficult. To explore further, we calculate the number of matchings throughout the \texttt{fabl} Gibbs sampler that violate our one-to-one assumption, and find on average 26.17 matchings in violation. These matchings are impossible under the \texttt{BRL} Gibbs sampler, so examining these more closely can provide insight how each model handles situations in which one file in $\bm{X}_1$ has
%multiple plausible matches in $\bm{X}_2$. 
%
%Table~\ref{Tab:rosa-maria} illustrates an example. Note that these records present a near violation of our assumption that there are no duplications within files; we continue to assume that records 825 and 826 in ERTL correspond to different individuals (most likely a mother and daughter), but their records are nearly identical. In addition, using the modified Levenstein distance, the comparison vectors \(\gamma_{2776, 825}\) and \(\gamma_{2776, 826}\) are exactly identical.
%
%\begin{table}
%	\centering
%	\begin{tabular}[h!]{llllllll}
%		\hline
%		Record No. & Dataset & Last Name & First Name & Day & Month & Year & Dept \\
%		\hline
%		825 & CDHES & Pineda & Rosa & 6 & 4 & 1984 & NA\\
%		826 & CDHES & Pineda & Rosa Maria & 6 & 4 & 1984 & NA\\
%		2776 & ERTL & Pineda & Rosa Maria & 4 & 4 & 1984 & Cuscatlan\\
%		\hline
%	\end{tabular}
%	\caption{Example of linkage situation in which record 2776 in ERTL has multiple plausible matches in CDHES, leading to undesirable behavior both \texttt{fabl} and \texttt{BRL}.}\label{Tab:rosa-maria}
%\end{table}
%
%Figure~\ref{fig:mixing-plot} shows the values that $Z_{825}$ and $Z_{826}$ take on throughout the Gibbs sampler, and demonstrates how each method handles this situation. Under \texttt{fabl}, both records in $\bm{X}_2$ match to the same record in $\bm{X}_1$ throughout the Gibbs process, creating consistent violations of one-to-one matching. Under \texttt{BRL}, the Gibbs process creates one matching configuration and stays there for a while. However, if one pair ``unmatches,'' then the other record has a chance to latch on. Then, the Gibbs process is stuck with that matching status for a while, resulting in a Gibbs process with poor mixing. 
%
%Through its independent sampling,\texttt{fabl} allows the modeler to inspect records with multiple plausible matches, and if they desire, to then choose the record pairing with the highest posterior probability. \texttt{BRL} in contrast, in strictly enforcing one-to-one matching throughout the sampler, can lead to situations where none of the plausible matches reach the threshold to be identified through the Bayes estimate. Thus relaxing the one-to-one constraint has served as a useful tool for identifying model misspecification. 
%
%\begin{figure}[h!]
%\begin{center}
%\includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
%\caption{Gibbs sampling in situation with multiple plausible matches. Under \texttt{BRL}, two records in ERTL compete for their most similar record, splitting the posterior probability. Under \texttt{fabl}, both record consistently match with their most similar record, consistently violating one-to-one matching.} \label{fig:mixing-plot}
%\end{center}
%\end{figure}

@@

%Tuning \texttt{fabl} for use in these scenarios is one potential avenue for
%future work.

%\begin{figure}[t]
%	
%	{\centering \includegraphics[width=0.6\textwidth]{../notes/figures/el_salvador/bad_mixing} 
%		
%	}
%	
%	\caption{Gibbs sampling in situation with multiple plausible matches.}\label{fig:mixing-plot}
%\end{figure}

%\subsection{National Long Term Care Survey}
%\label{nltcs} 
%
%The National Long Term Care Survey is a longitudinal study to track changes in health among Medicare recipients. It was conducted approximately every five years, and we concern ourselves here with linking records from years 1982 and 1989. The data only contains day, month, and year of birth, sex, state the individual lives in, and regional office. State and regional office have been anonymized, presented as numbers instead of their actual values. We construct the comparison vectors using only binary exact matching for each field, since there is no straightfoward way to incorporate partial matching in this context. The 1982 file contains $n_A = 20485$ records, 1989 file contains $n_B = 17466$ files, and the six linkage fields induces $2^6 = 64$ possible unique agreement patterns. We note that linkage quality is bound to be quite weak for many potential matches with such a large number of records and such limited identifying information.  
%
%With such large files, performing all-to-all comparisons under the original Fellegi-Sunter framework is infeasible. Indeed, this would require nearly 400 million comparison vectors, each containing 6 integer comparisons; the resulting data object would contain over 2 billion integers, and would require approximately 16 GBs of memory to hold. Instead, we take the partitioning approach reviewed in \ref{data-representation-hashing-and-storage}, and split each dataset into 10 smaller datasets. We use parallel computing on cluster service create the comparison vectors for each pair of data chunks, hash the results, and compress in size through storage efficient indexing. We then aggregate the results, creating a data object of just 87 MB, and proceed with the Gibbs Sampler. 
%
%Were a personal computer able to hold the 16 GB data object required by the direct approach, the Gibbs sampler under \texttt{BRL} would be tortuously slow, with computational complexity $O(n_A \times n_B \times F)$ where $n_A = 20485$. However, under \texttt{fabl} the sampler has complexity $O(P \times n_B \times F)$ where $P = 64$, so 1000 iterations of the Gibbs sampler are completed in just XX seconds. Posterior analysis yields an initial estimate of 9991 individuals linked across the the two datasets, and 8895 individuals linked once conflicts are resolved to ensure one-to-one matching. 
%
%Lastly, we comment on the significance of conducting this linkage without the use of deterministic blocking. With previous methods, one could link these two records by blocking on a certain field, and running separate Gibbs samplers on each of the separate blocks. This strategy drastically improves computation since it significantly reduces the number of comparison vectors that need to be formed and uses smaller data objects throughout the sampler. However, this computational savings comes at the cost of missing potential links, specifically in cases where the blocking variable is recorded in error. In this case, 62 of the matched pairs exhibit a disagreement in sex, 17 exhibit a disagreement in state, and 39 exhibit a disagreement in regional office. We have no way to verify that the matches found under \texttt{fabl} are indeed correct, so it is possible that matches found with such disagreements are false positives. Even if not, the modeler may conclude that, for example, 17 missed matches are a small price to pay to be able to block on state and run 50 smaller, independent linkage procedures. However, \texttt{fabl} allows the modeler to make this choice for themselves, rather than being forced to do so for computational reasons. Additionally, by running on the Gibbs sampler the entirety of both files, we achieve $m$ and $u$ parameters that are interpretale for the entire study, rather than separate parameters for each block. Although these $m$ and $u$'s are often regarded as nuisance parameters in the record linkage literature, \texttt{fabl} allows for study of these parameters that has been been possible before.

\bigskip

\bibliographystyle{jasa}
\bibliography{biblio}

>>>>

\documentclass[ba]{imsart}
%
\pubyear{0000}
\volume{00}
\issue{0}
\doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}
\newtheorem{lemma}{Lemma}
\usepackage{booktabs}       % professional-quality tables
\usepackage{algorithm}      % algorithm environment
\usepackage{algpseudocode}

\startlocaldefs
\input{subfiles/preamble.tex}
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}
%\title{A Sample Document\thanksref{T1}}
%\thankstext{T1}{<thanks text>}
\runtitle{Efficient and Scalable Bipartite Matching with Fast Beta Linkage  (fabl)}

\begin{aug}
\author{\fnms{Brian} \snm{Kundinger}\thanksref{addr1}\ead[label=e1]{brian.kundinger@duke.edu}},
\author{\fnms{Jerome} \snm{Reiter}\thanksref{addr1}\ead[label=e2]{jreiter@duke.edu}}
\and
\author{\fnms{Rebecca C.} \snm{Steorts}\thanksref{addr2}\ead[label=e3]{beka@stat.duke.edu}}

\runauthor{Kundinger et al.}

\address[addr1]{Department of Statistical Science,
  Duke University,
  P.O.\ Box 90251,
  Durham, NC 27708, USA
  \printead{e1}, % print email address of "e1"
  \printead*{e2}
}

\address[addr2]{Departments of Statistical Science and Computer Science,
  Duke University,
  P.O.\ Box 90251,
  Durham, NC 27708, USA
  \printead{e3}
}

%\thankstext{<id>}{<text>}
\end{aug}

\begin{abstract}
Recently, researchers have developed Bayesian versions of the Fellegi Sunter model for record linkage. These have the crucial advantage of quantifying uncertainty from imperfect linkages.  However, current implementations of Bayesian Fellegi Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks.  We propose and investigate a variation on Bayesian Fellegi Sunter models that we call fast beta linkage, or \texttt{fabl}.  Specifically, in \texttt{fabl} we propose independent prior distributions over the matching space, allowing us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large datasets through parallel computing and reduce memory costs through a new technique called storage efficient indexing. Through simulations and two case studies, we show that \texttt{fabl} has markedly increased speed with minimal loss of accuracy.
\end{abstract}

%% ** Keywords **
\begin{keyword}%[class=MSC]
\kwd{bipartite record linkage}
\kwd{Bayesian methods}
\kwd{hashing techniques}
\kwd{parallel/distributed computing}
\kwd{Markov chain Monte Carlo}
\end{keyword}

\end{frontmatter}

%% ** Mainmatter **

@@

@@

@@

%\section{}\label{}

% \begin{figure} 
% \includegraphics{<eps-file>}% place <eps-file> in ./img  subfolder
% \caption{}
% \label{}
% \end{figure}

% \begin{table} 
% *****************
% \begin{tabular}{lll}
% \end{tabular}
% *****************
% \caption{}
% \label{}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Supplementary Material, if any, should   %%
%% be provided in {supplement} environment  %%
%% with title and short description.        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \begin{supplement}
% \stitle{???}
% \sdescription{???.}
% \end{supplement}

\bibliographystyle{ba}
\bibliography{biblio}

% ** Acknowledgements **
% \begin{acknowledgement}
% \end{acknowledgement}

% ATTN: Should not include appendices in the main manuscript
%\input{subfiles/appendix.tex}
